---
title: "DSC 462: Assignment 4"
author: "Shouman Das"
date: "December 5, 2017"
output: pdf_document
header-includes:
  - \usepackage{amsmath,multirow}
  - \usepackage{tables}
  - \usepackage[caption=false]{subfig}
---
# Q1
(a)
$$
n = 2030, \quad \hat{p}= 1050/n
$$
$$
\begin{aligned}
CI_{95\%} &= \hat{p}\pm z_{0.025} \sqrt{\frac{\hat p(1-\hat p)}{n}}\\
&\approx 0.52 \pm 0.02
\end{aligned}
$$
```{r}
n = 2030
p = 1050/n
p
1.96* sqrt(p*(1-p)/n)
```
(b) Note that 
$$
\begin{aligned}
\alpha &= 0.05\\
Z_{\alpha/2} &= 1.96\\
E_o &= 0.01
\end{aligned}
$$
So we have $n = (\frac{Z_{\alpha/2}}{2E_o})^2 = 9604$.
```{r}
(1.96/2/0.01)^2
```
# Q2
(a) 
```{r}
n = 56
pHat = 20.5/n #for continuity correction
p0 = 0.25
Zobs = (pHat-p0)/sqrt(p0*(1-p0)/n)
alpha_obs = 1-pnorm(Zobs)
alpha_obs
```
So p-value = 0.0224 and we can reject $H_o$ at a significance level $\alpha = 0.05$.

(b) We can check the previous result by using R.
```{r}
prop.test(x = 21, n = 56, p = .25, alternative = "greater", correct = TRUE)
```
# Q3
We can do this problem by using R.
```{r}
n1 = 250
n2 = 500
x1 = 65
x2 = 100
prop.test(x =c(x2,x1), n = c(n2,n1), correct = F)
```
(1) So the 0.95 confidence interval for $p_2-p_1$ is $(-0.12469, 0.00469)$.

(2) Since the P-value is $0.0615$, the null hypothesis cannot be rejected.
\newpage

# Q4 
```{r}
n11 = 3
n12 = 31
n21 = 12
n22 = 45
OR = (n11*n22)/(n12*n21)
SE.log.OR = sqrt(1/n11+1/n12+1/n21+1/n22)
log(OR)
1.96*SE.log.OR
```
So 95% CI for the log odds ratis is $-1.10136 \pm 1.3453$. This also means that we cannot reject the null hypothesis, since $0$ is contained in the interval.

# Q5 
(a) To construct a 90% upper confidence bound for $\sigma$, we need 
$$
\chi^2_{n-1,1-\alpha} = \chi^2_{5,0.90} = 1.61
$$
which gives 
$$
\sigma < \frac{S_n}{\sqrt{\chi^2_{n-1,1-\alpha}/(n-1)}} = \frac{0.088}{\sqrt{1.61/5}} =  0.1550796
$$
```{r}
x = c(9.43, 9.85, 10.12, 9.89, 9.81, 10.3)
Sn = var(x)
Sn
Sn/sqrt(1.61/5)
```
(b) Using this upper bound for sample size estimation we get $n = 10$
```{r}
alpha = 0.05
z.alpha.half = qnorm(1-alpha/2)
E.o = 0.1
sigma = 0.1550796
(z.alpha.half*sigma/E.o)^2

```
\newpage

# Q6 
```{r}
n1 = 14
n2 = 18
S1.sq = 645.16
S2.sq = 1413.76
alpha = 0.1
F = S1.sq/S2.sq
crit.up = qf(1-0.05, 13,17)
crit.low = qf(0.05, 13,17)
F
crit.low
crit.up
```
So the rejection region is $F\leq 0.4$ or $F\geq 2.35$ So we cannot reject the null hypothesis. Also, we can get our p-value = 0.1569.
```{r}
2*min(pf(F, 13,17), 1- pf(F,13,17))

```

# Q7
First calculate the variance for null hypothesis.
$$
\sigma_0^2 = (1/2^2 + 1/6^2 + 1/3^2)\sigma_X^2 = 9.722
$$
Also the variance for the alternative hypothesis,
$$
\sigma_1^2 = \sigma_0^2+ 2/25*(0.56/12+0.18/6+0.21/18) = 14.13889
$$
So our hypothesis would be 
$$
H_{o}\ : \ \sigma^2 = \sigma_0^2, \quad H_{a} : \sigma^2 = \sigma_1^2
$$
Now our test statistic 
$$
W = \frac{(n-1)S^2_n}{\sigma_0^2}
$$
Now, if $H_o$ is true then $W \sim\chi^2_{n-1}$. If $H_a$ is true then $\frac{\sigma_0^2}{\sigma_1^2}W \sim \chi^2_{n-1}$. So our power would be
$$
\text{Power}(n) = P(\text{reject } H_0| H_a)
$$
$$
= P(W> \chi^2_{n-1,0.025}) = P(\frac{\sigma_0^2}{\sigma_1^2}\ W > \frac{\sigma_0^2}{\sigma_1^2}\ \chi^2_{n-1,0.025})
$$
where $\frac{\sigma_0^2}{\sigma_1^2} W \sim \chi^2_{n-1}$. 
```{r}
sigma0.sq = (1/4+1/36+1/9)*25
sigma1.sq = sigma0.sq + 2*25*(0.56/12+0.18/6+0.21/18)
Power = function(n){
    crit = sigma0.sq/sigma1.sq*qchisq(p = 0.05, df = n-1, lower.tail = FALSE)
    pchisq(q = crit, df = n-1, lower.tail = FALSE)
}
n = 2:200
plot(n, Power(n), type = 'l', main = "Power vs n")
abline(h = 0.9, col = 'blue')
n[Power(n)>=0.9][1]
```
We can see that 123 is the minimum number of sample required to have 90% power.
