---
title: "DSC 462: Assignment 4"
author: "Shouman Das"
date: "November 14, 2017"
output: pdf_document
header-includes:
  - \usepackage{amsmath,multirow}
  - \usepackage{tables}
  - \usepackage[caption=false]{subfig}
---

# Q1

(a) Standard deviation of the sample mean  $\sigma_{\bar{X}_{12}}=\dfrac{1.2}{\sqrt{12}}$ 
```{r}
1.2/sqrt(12)
```

(b) We have, $\bar{X}_{12} \sim N(76.5,1.2^2/12)$. The desired probability is then 
$$
\begin{aligned}
P(76.5-4< \bar{X}_{12} < 76.5+4) &= P(72.5<\bar{X}_{12}<80.5)\\
&= P(-4/0.3464102 <Z < 4/0.3464102)
&\approx 1
\end{aligned}
$$

# Q2
(a) Here the true variance of population is unknown. So we use $t_{n-1,\frac{\alpha}{2}}$ for $z_{\frac{\alpha}{2}}.$

```{r}
x = c(23.72, 18.95, 9.79, 12.04, 9.54)
t.test(x, conf.level = .9)
```
So 90% confidence interval is $(8.831717, 20.784283)$. We can also check this by hand calculation with $t_{4,.05}= 2.132$ and $\bar X_n = 14.808, S_n = 39.29347$, 
$$
(14.808- 2.132\sqrt{\frac{39.29347}{5}},14.808 + 2.132\sqrt{\frac{39.29347}{5}})
$$
$$
=(8.831717, 20.784283)
$$

\newpage
(b)
```{r}
xbar = mean(x)      # sample mean
mu0 = 24            # hypothetical value
s = sqrt(var(x))    # sample standard deviation 
n = 5               # sample size
Tobs = (xbar - mu0)/(s/sqrt(n)) 
Tobs                # test statistic
```

```{r}
alpha = 0.05
p_val = 2*pt(Tobs, df = n-1)
p_val
```
For significance level $\alpha = .05$ our P-value $\leq 0.05$, we have strong evidence to reject $H_0$ in favor of $H_a$. But for $\alpha = 0.01$, we don't have enough evidence to reject $H_0$.

# Q3 
```{r figs, fig.width=7,fig.height=6, fig.cap="\\label{fig:figs}Power curve"}
### set up labels with mathematical typesetting
ex0 = expression(paste("Power Curve with sample size ",
italic(n)," and ",alpha," = 0.05",sep=''))
ex1 = expression(italic(mu))
ex2 = expression(Power(mu))

### grid for horizontal axis
mu = seq(150, 175, by = 0.1)

### set up graphics window, draw empty plot (type='n')
par(mar=c(4,5,2,2), oma=c(4,4,4,4), cex=1,cex.axis=1.2, cex.lab=1.2, cex.main=1.2)
plot(range(mu), c(0,1), xlab=ex1, ylab=ex2, type='n')
title(ex0)

### grid
for (x in seq(150,175,by=2.5)) {abline(v=x,col='gray')}
for (y in seq(0,1,by=0.05)) {abline(h=y,col='gray')}
abline(h=0.05,col='blue')
ex4 = bquote(alpha == 0.05)
text(170, 0.05, ex4)

for(n in c(5,10,15,20,25,30)){
    # power curve for n
    alpha = 0.05
    z.alpha = qnorm(1-alpha)
    y = 1 - pnorm((150 + z.alpha*10/sqrt(n) -mu)/(10/sqrt(n)))
    lines(mu, y, type = 'l')
    
    # label n=5,30 plot
    if(n==5){
        ex3 = bquote(italic(n) == .(n))
        text(155, y[mu==155], ex3)
    }else if(n==30){
        ex3 = bquote(italic(n) == .(n))
        text(155, y[mu==155], ex3)
        }
}

```

(b) 
```{r, echo=FALSE}
library(xtable)
M = matrix(ncol = 6, nrow = 6)
Power = function(mu,n){
    alpha = 0.05
    z.alpha = qnorm(1-alpha)
    y = 1 - pnorm((150 + z.alpha*10/sqrt(n) -mu)/(10/sqrt(n)))
    return(y)
}
for(n in c(5,10,15,20,25,30)){
    for(mu in c(155,156,157,158,159,160)){
        M[n/5,mu-154] = Power(mu,n)
    }
    }
rownames(M) = paste("n = ", c(5,10,15,20,25,30))
colnames(M) = paste("mu =", c(155,156,157,158,159,160))
```


\begin{table}[ht]
\centering
$\mu$\\
\begin{tabular}{r|rrrrrr}
 
 \hline
 
 n&  155 &  156 & 157 &  158 &  159 &  160 \\ 
  \hline
 5 & 0.2992 & 0.3809 & 0.4683 & 0.5573 & 0.6434 & 0.7228 \\ 
   10 & 0.4746 & 0.5997 & 0.7152 & 0.8119 & 0.8852 & 0.9354 \\ 
   15 & 0.6147 & 0.7514 & 0.8568 & 0.9270 & 0.9672 & 0.9871 \\ 
   20 & 0.7228 & 0.8505 & 0.9313 & 0.9734 & 0.9913 & 0.9977 \\ 
   25 & 0.8038 & 0.9123 & 0.9682 & 0.9907 & 0.9978 & 0.9996 \\ 
   30 & 0.8630 & 0.9497 & 0.9857 & 0.9969 & 0.9995 & 0.9999 \\ 
   \hline
\end{tabular}
\end{table}

From the table, we can find the minimum sample size for 80% power. 

\begin{table}[ht]
\centering
\begin{tabular}{r|rrrrrr}
 \hline
 &  155 &  156 & 157 &  158 &  159 &  160 \\ 
  \hline
 \text{Min size for .8 power} & 25 & 20 & 15 & 10 & 10 & 10\\
   \hline
\end{tabular}
\end{table}

# Q4 
(a)i) Equal variance: 
```{r}
alpha = 0.02
xbar1 = 96.3
xbar2 = 121.5
S1 = 16.5
S2 = 12.5
n1 = 15
n2 = 8
Sp = sqrt(((n1-1)*S1^2+(n2-1)*S2^2)/(n1+n2-2))
t.crit = qt(1-alpha/2, df = n1+n2-2)
conf.int = c(xbar2-xbar1-t.crit*Sp*sqrt(1/n1+1/n2), xbar2-xbar1+t.crit*Sp*sqrt(1/n1+1/n2))
conf.int
```
ii) Unequal variance: Welch's procedure
```{r}
nu.w = floor((S1^2/n1+S2^2/n2)^2/((S1^2/n1)^2/(n1-1)+(S2^2/n2)^2/(n2-1)))
t.crit = qt(1-alpha/2, df = nu.w)
conf.int = c(xbar2-xbar1-t.crit*sqrt(S1^2/n1+S2^2/n2), xbar2-xbar1+t.crit*sqrt(S1^2/n1+S2^2/n2))
conf.int

```
(b) Hypothesis test for $H_o: \mu_2-\mu_1=0$ against $H_a:\mu_2-\mu_1>0$.
```{r}
# Equal variance
T.obs = (xbar2-xbar1)/(Sp*sqrt(1/n1+1/n2))
P.value = 1 - pt(T.obs, df = n1+n2-2)
P.value

# Unequal variance
T.obs = (xbar2-xbar1)/(sqrt(S1^2/n1+S2^2/n2))
P.value = 1 - pt(T.obs, df = nu.w)
P.value

```

In both cases, we can reject $H_o$ in favour of $H_a$.

# Q5
(a) t-test

```{r}
#paired t-test
pipe = c(6.64, 7.89, 1.83, 0.42, 0.85, 0.29, 0.57, 0.63, 0.32, 0.37, 0.00, 0.11, 4.86, 1.80, 0.23, 0.58)
brush = c(9.73, 8.21, 2.17, 0.75, 1.61, 0.75, 0.83, 0.56, 0.76, 0.32, 0.48, 0.52, 5.38, 2.33, 0.91, 0.79)
t.test(pipe, brush, paired = TRUE, conf.level = 0.99)
```
Since the p-value is less than $\alpha = 0.001$ we can reject the null hypothesis and say that brush is more effective than pipe. 

(b) 
```{r}
t.test(pipe, brush, var.equal = TRUE, conf.level = .99, paired = FALSE)
t.test(pipe, brush, var.equal = FALSE, conf.level = .99, paired = FALSE)
```
This we can see that $0$ is in the confidence interval. So we cannot reject $H_o$.

# Q5
(a) Note that $Var(Y) = \beta_1^2\ Var(X) + \beta_2^2\ Var(\epsilon) = \beta_1^2 + \beta_2^2$. So $\sigma_Y = \sqrt{\beta_1^2 + \beta_2^2}$. Also,
$cov(X,Y) =\ cov(X,\beta_1 \ X) = \beta_1$. So,
$$corr(X,Y) = \dfrac{\beta_1}{\sqrt{\beta_1^2+ \beta_2^2}}$$
which doesn't depend on $\beta_0$.

(b) We can choose $\beta_0 =0, \beta_1 = \rho, \beta_2 = \sqrt{1-\rho^2}$
\newpage
```{r}
set.seed(1234)
par(mfrow=c(2,2))
mat = matrix(nrow = 4, ncol = 4)
colnames(mat) <- c("rho", "Var(X)", "Var(Y)", "Correlation")
df  = as.data.frame(mat)
i = 1
for(rho in c(-0.5, 0, 0.5, 0.9)){
    x = rnorm(1000)
    epsilon = rnorm(1000)
    beta1 = rho
    beta2 = sqrt(1-rho^2)
    y = beta1*x+beta2*epsilon
    ex5 = bquote(rho == .(rho))
    plot(x,y, main = ex5, cex = 0.5)
    df$rho[i] = rho
    df$`Var(X)`[i] = var(x)
    df$`Var(Y)`[i] = var(y)
    df$Correlation[i] = cor(x,y)
    i=i+1
}
df # Summary of sample variance and correlation of X,Y
```

